Brother, Can You Spare a Manufacturing Job? How Voters React to Deindustrialization

### 2021-12-12 Final Draft/Final Draft.Rmd
updates to final draft

This paper analyzes the relationship between deindustrialization, race, and the election of Donald Trump in 2016. Previous analysis has  focused on racial animus, trade, or the conjunction of the two on  political polarization and election outcomes (\cite{Autor20}; \cite{Che16}; \cite{BR21}). We extend the analysis of \cite{Baccini21}, one of the first papers to examine the political effects of deindustrialization that while considering race and localization. Deindustrialization - the transition of the U.S. economy away from manufacturing - has led to widespread job loss and created adverse knock-on effects for a wide swath of Americans. Our measure of dindustrialization incorporates job losses due to trade with China and others (see \cite{Acemoglu16}); it also includes job losses due to automation, demand changes, and other factors. The experience of deindustrialization has not been uniform; rather, it has been concentrated in certain areas of the country that were previously economically reliant on manufacturing, as demonstrated in Figure \label{ManuMap}. The localized nature of this analysis highlights the uneven and concentrated harms of deindustrialization.

@@

\FloatBarrier
\begin{figure} \label{ManuMap}
\caption{Concentration of Deindustrialization (2004-2015) in the eastern Heartland}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center" , regfloat=FALSE}
# load map
model_input <- readRDS("../model_5_input.rds")

@@

ggplot() +
  geom_sf(data = mapping_county, aes(fill = ntile_mfg),
          color = NA, size = .01) +
  scale_fill_manual(values = colors,
                    labels = labels) +
  # add state lines
  geom_sf(data = state_shape, fill = NA, color = "black", size = .1) +
  coord_sf(datum = NA) +
  theme_minimal() +
  # geom_sf_text_repel(aes(label = state), data = mapping_small,
  #                    force = .00001,  seed = 10,
  #                    size = 2.5) +
  theme(  axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          text = element_text(size = 8)) +
  labs(title = "Net Change in Manufacturing Employment from 2004 to 2015",
       subtitle = "",
       fill = "") +
  theme(legend.key.size = unit(.4, 'cm'))

\end{figure}
\FloatBarrier

Like \cite{Baccini21}, we find evidence that the experience of deindustralization was associated with support for Donald Trump, particularly among whites. \cite{Baccini21}
find that gross manufacturing layoffs between 2012-2015 were associated with a loss in Democratic vote share. However, using a different measure of deindustrialization - *net* change in manufacturing employment - we find the oppposite effect: that counties with manufacturing job *gains* from 2012-2015 tended to swing towards Donald Trump in 2016. We extend the analysis to the 2004-2015 period in order to capture the long-term effects of deindustrialization on a region. Doing so, we find that regions experiencing more manufacturing job losses swung towards Donald Trump in 2016, and that this occurred more when those losing jobs were white.

Our analysis extends \cite{Baccini21} in several ways. First, we change focus from a measure of gross losses in manufacturing jobs to net losses. As discussed in \ref{datasec}, we believe this to be a more accurate measure of deindustrialization. Second, we change the temporal focus. Rather than focusing on manufacturing job loss in 2012-2015, we extend our analysis back to 2004. This is an important change. As shown in Figure \ref{natlPlot}, the bulk of U.S. manufacturing job loss took place prior to 2012; in fact, the 2012-2015 period was characterized by a slight rebound in manufacturing jobs. During this rebound, counties that experienced the most pronounced manufacturing job losses in prior years were also likely to experience the most pronounced recovery, as described in Figure \ref{vShapePlot}. Because of this v-shaped recovery, focusing only on 2012-2015 job losses/gains paints a distorted picture. Finally, we augment the instrumental-variable approach with one based on covariate balancing, which produces similar results.

<!-- FIGURE X: TIME SERIES OF MANUFACUTRING EMPLOYMENT  -->
\FloatBarrier
\begin{figure} \label{natlPlot}
\caption{Nationwide Change in Manufacturing Employment, 1995-2019}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}
# manufacturing layoffs over time

@@

```
\begin{figurenotes}
"Gain or loss of national manufacturing jobs, divided by total national workforce."
\end{figurenotes}
\end{figure}
\FloatBarrier

@@

Following \cite{Baccini21}, our unit of analysis is the county. This allows us to capture not only the direct effects of deindustrialization on laid-off manufacturing workers, but also on the local economy. The primary outcome variable is the county's change in Democratic vote share between 2012 and 2016. We obtain this from 
<!--  sources of electoral data -->

We want to measure the causal relationship between deindustrialization and the change in vote share. As a measure of deindustrialization, we use the loss of manufacturing jobs as a share of the total beginning-of-period employment in each county. Suppose at the beginning of our period there were $2000$ manufacturing workers in a county, and $8000$ non-manufacturing workers. At the end there are only $1500$ manufacturing workers. This is a loss of $500/10000 = 5\%.$

@@

\FloatBarrier
\begin{figure} \label{counterfactual2}
\caption{Counterfactual Presidential Election Simulation Results}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}

>>>>

\section{Introduction} 
\label{Introduction}

This paper analyzes the relationship between deindustrialization, race, and the election of Donald Trump in 2016. Previous analyses have  focused on the effects of racial animus, trade, or the conjunction of the two on  political polarization and election outcomes (\cite{Autor20}; \cite{Che16}; \cite{BR21}). We extend the analysis of \cite{Baccini21}, one of the first papers to examine the political effects of deindustrialization while considering race and localization.

Deindustrialization - the transition of the U.S. economy away from manufacturing - has led to widespread job loss and created adverse knock-on effects for a wide swath of Americans. Our measure of dindustrialization incorporates job losses due to trade with China and others (see \cite{Acemoglu16}); it also includes job losses due to automation, demand changes, and other factors. The experience of deindustrialization has not been uniform; rather, it has been concentrated in areas of the country that were previously economically reliant on manufacturing. As shown in Figure \ref{ManuMap}, states across the eastern half of America, particularly in the South and Midwest, have experienced the most pronouncd manufacturing employment losses. The localized nature of this analysis highlights the uneven and concentrated harms of deindustrialization.

@@

\FloatBarrier
\begin{figure} \label{ManuMap}
\caption{Concentration of Deindustrialization (2004-2015) in the eastern United States }
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center" , regfloat=FALSE}
# load map
model_input <- readRDS("../model_5_input.rds")

@@

ggplot() +
  geom_sf(data = mapping_county, aes(fill = ntile_mfg),
          color = NA, size = .01) +
  scale_fill_manual(values = colors,
                    labels = labels) +
  # add state lines
  geom_sf(data = state_shape, fill = NA, color = "black", size = .1) +
  coord_sf(datum = NA) +
  theme_minimal() +
  # geom_sf_text_repel(aes(label = state), data = mapping_small,
  #                    force = .00001,  seed = 10,
  #                    size = 2.5) +
  theme(  axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          text = element_text(size = 8)) +
  labs(title = "",
       subtitle = "",
       fill = "Net Manufacturing Employment\nLosses from 2004 to 2015") +
  theme(legend.key.size = unit(.4, 'cm'))

```
\begin{figurenotes}
Calculated at the county-level as the net change in manufacturing employment between 2004 and 2015 divided by 2004 total employment.
\end{figurenotes}
\end{figure}
\FloatBarrier

In their central analysis, \cite{Baccini21} analyzed the effect of gross manufacturing job losses from 2012 to 2015 on the change in Democratic vote share in the 2016 presidential election. \cite{Baccini21} find evidence that the experience of deindustralization was associated with support for Donald Trump, particularly among whites.

Our analysis extends \cite{Baccini21} in several ways. First, we change focus from a measure of *gross* losses in manufacturing jobs to *net* losses. As discussed in \ref{datasec}, we believe this to be a more accurate measure of deindustrialization. Second, we change the temporal focus. Rather than focusing on manufacturing job loss in 2012-2015, we extend our analysis back to 2004. This is an important change. As shown in Figure \ref{natlPlotTS}, the bulk of U.S. manufacturing job loss took place prior to 2012; in fact, the 2012-2015 period was characterized by a slight rebound in manufacturing jobs. Finally, we augment the instrumental-variable approach with one based on covariate balancing.

Like \cite{Baccini21}, our results suggest that the experience of deindustralization was associated with support for Donald Trump, particularly among whites. However, our results differ from \cite{Baccini21} in regards to the *timing* of the impact of manufacturing job losses on employment. In fact, using our preferred measure of deindustrialization - *net* change in manufacturing employment - we find that counties with manufacturing job *gains* from 2012-2015 tended to swing towards Donald Trump in 2016. Given this counter-intuitive result, we extend the analysis to the 2004-2015 period in order to capture the long-term effects of deindustrialization on a region. Doing so, we find that regions experiencing more manufacturing job losses from 2004-2015 swung towards Donald Trump in 2016, and that this occurred more when those losing jobs were white. We get similar result using our covariate-matching approach. Relative to \cite{Baccini21}, our results point to the substantive importance of the timing of manufacturing layoffs on political outcomes.

<!-- FIGURE X: TIME SERIES OF MANUFACUTRING EMPLOYMENT  -->
\FloatBarrier
\begin{figure} \label{natlPlotTS}
\caption{Nationwide Change in Manufacturing Employment, 1995-2019}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}
# manufacturing layoffs over time

@@

```
\begin{figurenotes}
Annual net change in manufacturing employment in the United States. 
\end{figurenotes}
\end{figure}
\FloatBarrier

@@

Following \cite{Baccini21}, our unit of analysis is the county. This allows us to capture not only the direct effects of deindustrialization on laid-off manufacturing workers, but also possible spillover effects on the greater population within the county. The primary outcome variable is the county's change in Democratic vote share between 2012 and 2016. We obtain this from the publically-available replication data provided by \cite{Baccini21}.

We want to measure the causal relationship between deindustrialization and the change in vote share. As a measure of deindustrialization, we use the loss of manufacturing jobs as a share of the total beginning-of-period employment in each county. Suppose at the beginning of our period there were $2000$ manufacturing workers in a county, and $8000$ non-manufacturing workers. At the end there are only $1500$ manufacturing workers. Using our measure of net change in manufacturing employment, this is a loss of $500/10000 = 5\%.$

@@

\FloatBarrier
\begin{figure} \label{counterfactual2}

\caption{Clinton Electoral College Votes}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

### 2021-12-12 Final Draft/Final-Draft.tex
adding placeins package
updates to final draft

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

This paper analyzes the relationship between deindustrialization, race,
and the election of Donald Trump in 2016. Previous analysis has focused
on racial animus, trade, or the conjunction of the two on political
polarization and election outcomes (\cite{Autor20}; \cite{Che16};
\cite{BR21}). We extend the analysis of \cite{Baccini21}, one of the
first papers to examine the political effects of deindustrialization
that while considering race and localization. Deindustrialization - the
transition of the U.S. economy away from manufacturing - has led to
widespread job loss and created adverse knock-on effects for a wide
swath of Americans. Our measure of dindustrialization incorporates job
losses due to trade with China and others (see \cite{Acemoglu16}); it
also includes job losses due to automation, demand changes, and other
factors. The experience of deindustrialization has not been uniform;
rather, it has been concentrated in certain areas of the country that
were previously economically reliant on manufacturing, as demonstrated
in Figure \label{ManuMap}. The localized nature of this analysis
highlights the uneven and concentrated harms of deindustrialization.

\begin{figure} \label{ManuMap}
\caption{Concentration of Deindustrialization (2004-2015) in the eastern Heartland}

@@

\end{figure}

Like \cite{Baccini21}, we find evidence that the experience of
deindustralization was associated with support for Donald Trump,
particularly among whites. \cite{Baccini21} find that gross
manufacturing layoffs between 2012-2015 were associated with a loss in
Democratic vote share. However, using a different measure of
deindustrialization - \emph{net} change in manufacturing employment - we
find the oppposite effect: that counties with manufacturing job
\emph{gains} from 2012-2015 tended to swing towards Donald Trump in
2016. We extend the analysis to the 2004-2015 period in order to capture
the long-term effects of deindustrialization on a region. Doing so, we
find that regions experiencing more manufacturing job losses swung
towards Donald Trump in 2016, and that this occurred more when those
losing jobs were white.

Our analysis extends \cite{Baccini21} in several ways. First, we change
focus from a measure of gross losses in manufacturing jobs to net
losses. As discussed in \ref{datasec}, we believe this to be a more
accurate measure of deindustrialization. Second, we change the temporal
focus. Rather than focusing on manufacturing job loss in 2012-2015, we
extend our analysis back to 2004. This is an important change. As shown
in Figure \ref{natlPlot}, the bulk of U.S. manufacturing job loss took
place prior to 2012; in fact, the 2012-2015 period was characterized by
a slight rebound in manufacturing jobs. During this rebound, counties
that experienced the most pronounced manufacturing job losses in prior
years were also likely to experience the most pronounced recovery, as
described in Figure \ref{vShapePlot}. Because of this v-shaped recovery,
focusing only on 2012-2015 job losses/gains paints a distorted picture.
Finally, we augment the instrumental-variable approach with one based on
covariate balancing, which produces similar results.

\begin{figure} \label{natlPlot}
\caption{Nationwide Change in Manufacturing Employment, 1995-2019}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-3-1} \end{center}
\begin{figurenotes}
"Gain or loss of national manufacturing jobs, divided by total national workforce."
\end{figurenotes}
\end{figure}

@@

Following \cite{Baccini21}, our unit of analysis is the county. This
allows us to capture not only the direct effects of deindustrialization
on laid-off manufacturing workers, but also on the local economy. The
primary outcome variable is the county's change in Democratic vote share
between 2012 and 2016. We obtain this from XXXXXXXXX

We want to measure the causal relationship between deindustrialization
and the change in vote share. As a measure of deindustrialization, we
use the loss of manufacturing jobs as a share of the total
beginning-of-period employment in each county. Suppose at the beginning
of our period there were \(2000\) manufacturing workers in a county, and
\(8000\) non-manufacturing workers. At the end there are only \(1500\)
manufacturing workers. This is a loss of \(500/10000 = 5\%.\)

@@

XXXXXXXXXXXXXXXX sources of other data

Basic descriptive statistics of job losses and other key covariates can
be seen in Table \ref{tab:descTable}.

@@

\caption{\label{tab:descTable}Manufacturing Job Changes 2004-2015}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrr}
\toprule
  & Mean & Std Dev & 25th Pctile & Median & 75th Pctile\\
\midrule
Mfg Share of Emp & 0.20 & 0.15 & 0.09 & 0.17 & 0.29\\
Mfg Share of Emp (White) & 0.16 & 0.13 & 0.06 & 0.13 & 0.22\\
Mfg Share of Emp (Nonwhite) & 0.05 & 0.07 & 0.01 & 0.02 & 0.06\\
Change in Mfg Jobs/ Worker & 0.04 & 0.09 & -0.01 & 0.02 & 0.06\\
Change in Mfg Jobs/ Worker (W) & 0.03 & 0.07 & 0.00 & 0.02 & 0.06\\
\addlinespace
Change in Mfg Jobs/ Worker (NW) & 0.00 & 0.04 & -0.01 & 0.00 & 0.01\\
\bottomrule
\end{tabular}}
\end{table}

@@

The results can be found in Table \ref{regResult04}. The regression
coefficient in Model 4, for instance, means that a loss of \(10\)
manufacturing jobs per \(100\) workers in a county (a relatively extreme
value) corresponds to a loss of \(3.2 \%\) in Democratic vote share.

\begin{table}[!htbp] \centering 
  \caption{Effect of Manufacturing Job Loss from 2004 to 2015} 
  \label{regResult04} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{Change in Dem. Share, 2016-2012} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 Manufacturing Layoffs & $-$0.34$^{**}$ & $-$0.38$^{**}$ &  &  \\ 
  & (0.14) & (0.17) &  &  \\ 
  & & & & \\ 
 White Manufacturing Layoffs &  &  & $-$0.23$^{***}$ & $-$0.32$^{***}$ \\ 
  &  &  & (0.08) & (0.10) \\ 
  & & & & \\ 
 Nonwhite Manufacturing Layoffs &  &  & 0.15$^{***}$ & 0.20$^{***}$ \\ 
  &  &  & (0.05) & (0.06) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
White Share/Svc Layoffs Control & No & Yes & No & Yes \\ 
Observations & 3,049 & 3,049 & 2,750 & 2,750 \\ 
Adjusted R$^{2}$ & 0.72 & 0.72 & 0.75 & 0.75 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

This coefficient is still somewhat difficult to interpret, and made
increasingly so by the winner-take-all system by which U.S. Electoral
College votes are allocated. As an alternative perspective, we examined
a range of counterfactual scenarios, in which counties experienced job
losses XXXXXXXXXXXXXXXXXXXXXX EXPLAIN COUNTERFACTUAL

\begin{figure} \label{counterfactual1}
\caption{Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-6-1} \end{center}
\end{figure}

\begin{figure} \label{counterfactual2}
\caption{Clinton Electoral College Votes}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-7-1} \end{center}
\end{figure}

\begin{figure} \label{counterfactual3}
\caption{Clinton Winning Freq}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-8-1} \end{center}
\end{figure}

\subsubsection{Timing}

Restricting our analysis to only the 2012-2015 period, while continuing
to use the net job loss metric, produces an effect opposite to that of
\cite{Baccini21}. The loss of Democratic vote share seems to be
associated with a gain in manufacturing jobs, both in general and among
whites. We believe this time-restricted analysis suffers from an omitted
variable bias. As discussed above, the 2012-2015 period can be broadly
characterized as a slight rebound from the manufacturing job losses in
the 2008 financial crisis. The counties that experience job losses from
2004-2012 are more likely to have job gains in 2012-2015 (see
\ref{vShapePlot}). An apparent relationship between 2012-2015 job gains
and loss of Democratic vote share in fact represents 2004-2012 job
losses that may lead to both the later job gains and the loss of vote
share. These results can be seen in Table \ref{regResult12}.

\begin{table}[!htbp] \centering 
  \caption{Effect of Manufacturing Job Loss from 2004 to 2015} 
  \label{regResult12} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{Change in Dem. Share, 2016-2012} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 Manufacturing Layoffs & 0.43$^{**}$ & 0.37$^{**}$ &  &  \\ 
  & (0.17) & (0.17) &  &  \\ 
  & & & & \\ 
 White Manufacturing Layoffs &  &  & 6.50$^{***}$ & 2.33$^{***}$ \\ 
  &  &  & (2.20) & (0.76) \\ 
  & & & & \\ 
 Nonwhite Manufacturing Layoffs &  &  & $-$3.99$^{***}$ & $-$1.42$^{***}$ \\ 
  &  &  & (1.36) & (0.47) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
White Share/Svc Layoffs Control & No & Yes & No & Yes \\ 
Observations & 3,064 & 3,064 & 2,765 & 2,765 \\ 
Adjusted R$^{2}$ & 0.73 & 0.73 & 0.75 & 0.75 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

@@

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-10-1} \end{center}
\begin{figurenotes}
"Stuff here"
\end{figurenotes}
\end{figure}

@@

>>>>

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{placeins}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\section{Introduction} 
\label{Introduction}

This paper analyzes the relationship between deindustrialization, race,
and the election of Donald Trump in 2016. Previous analyses have focused
on the effects of racial animus, trade, or the conjunction of the two on
political polarization and election outcomes (\cite{Autor20};
\cite{Che16}; \cite{BR21}). We extend the analysis of \cite{Baccini21},
one of the first papers to examine the political effects of
deindustrialization while considering race and localization.

Deindustrialization - the transition of the U.S. economy away from
manufacturing - has led to widespread job loss and created adverse
knock-on effects for a wide swath of Americans. Our measure of
dindustrialization incorporates job losses due to trade with China and
others (see \cite{Acemoglu16}); it also includes job losses due to
automation, demand changes, and other factors. The experience of
deindustrialization has not been uniform; rather, it has been
concentrated in areas of the country that were previously economically
reliant on manufacturing. As shown in Figure \ref{ManuMap}, states
across the eastern half of America, particularly in the South and
Midwest, have experienced the most pronouncd manufacturing employment
losses. The localized nature of this analysis highlights the uneven and
concentrated harms of deindustrialization.

\FloatBarrier
\begin{figure} \label{ManuMap}
\caption{Concentration of Deindustrialization (2004-2015) in the eastern United States }

@@

\FloatBarrier
\begin{figurenotes}
Calculated at the county-level as the net change in manufacturing employment between 2004 and 2015 divided by 2004 total employment.
\end{figurenotes}
\end{figure}
\FloatBarrier

In their central analysis, \cite{Baccini21} analyzed the effect of gross
manufacturing job losses from 2012 to 2015 on the change in Democratic
vote share in the 2016 presidential election. \cite{Baccini21} find
evidence that the experience of deindustralization was associated with
support for Donald Trump, particularly among whites.

Our analysis extends \cite{Baccini21} in several ways. First, we change
focus from a measure of \emph{gross} losses in manufacturing jobs to
\emph{net} losses. As discussed in \ref{datasec}, we believe this to be
a more accurate measure of deindustrialization. Second, we change the
temporal focus. Rather than focusing on manufacturing job loss in
2012-2015, we extend our analysis back to 2004. This is an important
change. As shown in Figure \ref{natlPlotTS}, the bulk of U.S.
manufacturing job loss took place prior to 2012; in fact, the 2012-2015
period was characterized by a slight rebound in manufacturing jobs.
Finally, we augment the instrumental-variable approach with one based on
covariate balancing.

Like \cite{Baccini21}, our results suggest that the experience of
deindustralization was associated with support for Donald Trump,
particularly among whites. However, our results differ from
\cite{Baccini21} in regards to the \emph{timing} of the impact of
manufacturing job losses on employment. In fact, using our preferred
measure of deindustrialization - \emph{net} change in manufacturing
employment - we find that counties with manufacturing job \emph{gains}
from 2012-2015 tended to swing towards Donald Trump in 2016. Given this
counter-intuitive result, we extend the analysis to the 2004-2015 period
in order to capture the long-term effects of deindustrialization on a
region. Doing so, we find that regions experiencing more manufacturing
job losses from 2004-2015 swung towards Donald Trump in 2016, and that
this occurred more when those losing jobs were white. We get similar
result using our covariate-matching approach. Relative to
\cite{Baccini21}, our results point to the substantive importance of the
timing of manufacturing layoffs on political outcomes.

\FloatBarrier
\begin{figure} \label{natlPlotTS}
\caption{Nationwide Change in Manufacturing Employment, 1995-2019}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-3-1} \end{center}

\FloatBarrier
\begin{figurenotes}
Annual net change in manufacturing employment in the United States. 
\end{figurenotes}
\end{figure}
\FloatBarrier

@@

Following \cite{Baccini21}, our unit of analysis is the county. This
allows us to capture not only the direct effects of deindustrialization
on laid-off manufacturing workers, but also possible spillover effects
on the greater population within the county. The primary outcome
variable is the county's change in Democratic vote share between 2012
and 2016. We obtain this from the publically-available replication data
provided by \cite{Baccini21}.

We want to measure the causal relationship between deindustrialization
and the change in vote share. As a measure of deindustrialization, we
use the loss of manufacturing jobs as a share of the total
beginning-of-period employment in each county. Suppose at the beginning
of our period there were \(2000\) manufacturing workers in a county, and
\(8000\) non-manufacturing workers. At the end there are only \(1500\)
manufacturing workers. Using our measure of net change in manufacturing
employment, this is a loss of \(500/10000 = 5\%.\)

@@

Basic descriptive statistics of job losses and other key covariates can
be seen in Table \ref{tab:descTable}. \FloatBarrier

@@

\caption{\label{tab:descTable}Manufacturing Job Changes 2004-2015}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrr}
\toprule
  & Mean & Std Dev & 25th Pctile & Median & 75th Pctile\\
\midrule
Mfg Share of Emp & 0.20 & 0.15 & 0.09 & 0.17 & 0.29\\
Mfg Share of Emp (White) & 0.16 & 0.13 & 0.06 & 0.13 & 0.22\\
Mfg Share of Emp (Nonwhite) & 0.05 & 0.07 & 0.01 & 0.02 & 0.06\\
Change in Mfg Jobs/ Worker & 0.04 & 0.09 & -0.01 & 0.02 & 0.06\\
Change in Mfg Jobs/ Worker (W) & 0.03 & 0.07 & 0.00 & 0.02 & 0.06\\
\addlinespace
Change in Mfg Jobs/ Worker (NW) & 0.00 & 0.04 & -0.01 & 0.00 & 0.01\\
\bottomrule
\end{tabular}}
\end{table}
\FloatBarrier

@@

The results can be found in Table \ref{regResult04}. The regression
coefficient in Model 4, for instance, means that a loss of \(10\)
manufacturing jobs per \(100\) workers in a county (a relatively extreme
value) corresponds to a loss of \(3.2 \%\) in Democratic vote share.
\FloatBarrier

\begin{table}[!htbp] \centering 
  \caption{Effect of Manufacturing Job Loss from 2004 to 2015} 
  \label{regResult04} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{Change in Dem. Share, 2016-2012} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 Manufacturing Layoffs & $-$0.34$^{**}$ & $-$0.38$^{**}$ &  &  \\ 
  & (0.14) & (0.17) &  &  \\ 
  & & & & \\ 
 White Manufacturing Layoffs &  &  & $-$0.23$^{***}$ & $-$0.32$^{***}$ \\ 
  &  &  & (0.08) & (0.10) \\ 
  & & & & \\ 
 Nonwhite Manufacturing Layoffs &  &  & 0.15$^{***}$ & 0.20$^{***}$ \\ 
  &  &  & (0.05) & (0.06) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
White Share/Svc Layoffs Control & No & Yes & No & Yes \\ 
Observations & 3,049 & 3,049 & 2,750 & 2,750 \\ 
Adjusted R$^{2}$ & 0.72 & 0.72 & 0.75 & 0.75 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table} 
\FloatBarrier

This coefficient is still somewhat difficult to interpret, and made
increasingly so by the winner-take-all system by which U.S. Electoral
College votes are allocated. As an alternative perspective, we examined
a range of counterfactual scenarios, in which counties experienced job
losses

\FloatBarrier
\begin{figure} \label{counterfactual1}
\caption{Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-6-1} \end{center}

\FloatBarrier
\end{figure}
\FloatBarrier

\FloatBarrier
\begin{figure} \label{counterfactual2}

\caption{Clinton Electoral College Votes}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-7-1} \end{center}

\FloatBarrier
\end{figure}
\FloatBarrier

\FloatBarrier
\begin{figure} \label{counterfactual3}
\caption{Clinton Winning Freq}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-8-1} \end{center}

\FloatBarrier
\end{figure}
\FloatBarrier
\subsubsection{Timing}

Restricting our analysis to only the 2012-2015 period, while continuing
to use the net job loss metric, produces an effect opposite to that of
\cite{Baccini21}. The loss of Democratic vote share seems to be
associated with a gain in manufacturing jobs, both in general and among
whites. We believe this time-restricted analysis suffers from an omitted
variable bias. As discussed above, the 2012-2015 period can be broadly
characterized as a slight rebound from the manufacturing job losses in
the 2008 financial crisis. The counties that experience job losses from
2004-2012 are more likely to have job gains in 2012-2015 (see
\ref{vShapePlot}). An apparent relationship between 2012-2015 job gains
and loss of Democratic vote share in fact represents 2004-2012 job
losses that may lead to both the later job gains and the loss of vote
share. These results can be seen in Table \ref{regResult12}.
\FloatBarrier

\begin{table}[!htbp] \centering 
  \caption{Effect of Manufacturing Job Loss from 2004 to 2015} 
  \label{regResult12} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{4}{c}{Change in Dem. Share, 2016-2012} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 Manufacturing Layoffs & 0.43$^{**}$ & 0.37$^{**}$ &  &  \\ 
  & (0.17) & (0.17) &  &  \\ 
  & & & & \\ 
 White Manufacturing Layoffs &  &  & 6.50$^{***}$ & 2.33$^{***}$ \\ 
  &  &  & (2.20) & (0.76) \\ 
  & & & & \\ 
 Nonwhite Manufacturing Layoffs &  &  & $-$3.99$^{***}$ & $-$1.42$^{***}$ \\ 
  &  &  & (1.36) & (0.47) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
White Share/Svc Layoffs Control & No & Yes & No & Yes \\ 
Observations & 3,064 & 3,064 & 2,765 & 2,765 \\ 
Adjusted R$^{2}$ & 0.73 & 0.73 & 0.75 & 0.75 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table} 
\FloatBarrier

@@

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-10-1} \end{center}

\FloatBarrier
\begin{figurenotes}
"Stuff here"
\end{figurenotes}
\end{figure}

@@

\FloatBarrier
